---
title: "Assignment 2R"
author: "Benjamin Moran"
date: "16 August 2016"
output: html_document
---
```{r setup, include = FALSE}
require(astsa)
require(ggplot2)
```
### Assignment 2: R Coding
Submission: You should email your R-code to ‘Ali.Eshragh@newcastle.edu.au’ by Friday, August 26, 11:55pm. Write your full name and student ID No. on the top of your code and your file name must be your complete surname.

The quarterly earnings per share for the U.S. company Johnson & Johnson, 1960-1984, are stored in the variable $jj$ from the R-package astsa. Write an R-code to address the following questions:

#####1. How many data are stored in the variable jj?
```{r Q1}
length(jj)
```

#####2. What are the average and variance of the Johnson & Johnson quarterly earnings per share from 1960 to 1984?

```{r Q2}
mean(jj)
var(jj)
var(jj)*(length(jj)/(length(jj)-1))
```

#####3. Plot the time series of the Johnson & Johnson quarterly earnings per share from 1960 to 1984. The labels of the horizontal and vertical axes should be in bold-style with the titles ‘Quarterly Earnings per Share’ and ‘Time’, respectively.

```{r Q3}
plot.ts(
  jj,
  xlab=expression(
    bold("Quarterly Earnings per Share")
    ),
  ylab=expression(
    bold("Time")
    ),
  main="Johnson & Johnson Quarterly Earnings per Share: \n 1960 - 1984"
  )

```

#####4. Plot the autocovariance function of this time series within 16 lags. The main title of this plot should be ‘Johnson & Johnson Quarterly Earnings per Share’.

```{r Q4}

acf(jj, 16, type=c("covariance"), main="Johnson & Johnson Quarterly Earnings per Share")

acf(jj, 16, type=c("correlation"), main="Johnson & Johnson Quarterly Earnings per Share")

acf(jj, 16, type=c("partial"), main="Johnson & Johnson Quarterly Earnings per Share")

```

####References
R.H. Shumway and D.S. Stoffer, Time Series Analysis and Its Applications With R Examples,
Springer, New York, 2010.
