---
title: 'Assignment 1: R Coding'
author: "Benjamin G. Moran, c3076448@uon.edu.au"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    includes: 
      before_body: eqnnumber.js
    mathjax: default
---
***
1. Define the $1000 × 2$ matrix $z$.

```{r Question 1}
z <- matrix(nrow = 1000, ncol = 2)
summary(z)
```
***
2. Generate two sets of $1000$ Standard Normal random variates by using equations $z1$ and $z2$ given
in item $20$ and store each of them in one column of matrix $z$.

```{r Question 2, results='hide'}
set.seed(2936)
u1 <- runif(1000,0,1)
u2 <- runif(1000,0,1)
z1 = sqrt(-2*log(u1))*cos(2*pi*u2)
z2 = sqrt(-2*log(u1))*sin(2*pi*u2)

z[,1] <- z1
z[,2] <- z2
  
```

```{r Question 2 Output}
summary(z)
```
***
3. Test, statistically, if each column of matrix $z$ is normally distributed at the significance level of
$5\%$.
```{r Normality Test}
normtest.z1 <- broom::tidy(shapiro.test(z[,1]))
normtest.z2 <- broom::tidy(shapiro.test(z[,2]))
```

```{r, echo=FALSE}
z1.p <- normtest.z1$p.value
if(z1.p > 0.05) {
  print(paste("z1 p-value =", z1.p, " - therefore, z1 is normally distributed"))
} else{
    print(paste("z1 p-value =", z1.p, " - therefore, z1 is not normally distributed"))
  }
z2.p <- normtest.z2$p.value
if(normtest.z2$p.value > 0.05) {
  print(paste("z2 p-value =", z2.p, " - therefore, z2 is normally distributed"))
} else{
    print(paste("z2 p-value =", z2.p, " - therefore, z2 is not normally distributed"))
  }
```
***
4. Test, statistically, if two columns of matrix $z$ are uncorrelated at the significance level of $5\%$.
```{r Correlation Test}
cortest.z <- broom::tidy(cor.test(z[,1],z[,2]))
cortest.z$p.value
```
***
5. Store both columns of matrix $z$ in a long vector, say $v$.

```{r}
v <- c(z[,1], z[,2])
summary(v)
length(v)

```
***
6. Find the average and variance of the generated data in $v$.

```{r}
mean(v)
var(v)
```
***
7. Define a new vector   $w = 2v + 5$. Show, numerically, that the average of the sample data in $w$ is equal to two time the average of the sample data in $v + 5$, and the variance of the sample data in $w$ is four times the variance of the sample data in $v$.

```{r }
w <-  2*v + 5
summary(w)

avg.w <- mean(w)
avg.v <- mean(v)

diff.avg <- 2*avg.v + 5 - avg.w

diff.avg

var.w <- var(w)
var.v <- var(v)

diff.var <- 4*var.v - var.w

diff.var

```