---
title: "Lab1"
author: "Benjamin Moran"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include = FALSE}
```
##Introduction to R

1. Help:

```{r 1, eval = FALSE} 
help(boxplot) 

#or 

?boxplot
```

2. Comment:

```{r 2, eval = FALSE}
# Your comment . . .
```

3. Get the current working directory:

```{r 3}
getwd()
```

4. Change the current working directory:

```{r 4, eval = FALSE}
setwd("c:/")
```

5. Define a variable:

```{r 5}
a <- 10 

#or

a = 10
```

6. Define a vector:

```{r 6} 
ID <- c(11, 12, 13, 14, 15) 
G <- c(81, 53, 66, 72, 68) 

#‘c’ stands for ‘concatenate’.
```

7. View the of elements of a vector:

```{r 7}
ID
ID[2:4]
ID[-2]
```

8. Find the number of elements in a vector:

```{r 8}
length(ID)
```

9. Find the summation, average, variance, minimum and maximum of a vector:

```{r 9}
sum(G)
mean(G)
var(G)
min(G)
max(G)
```

10. Combine variables/vectors:

```{r 10}
IDG <- c(ID,G)
CL1 <- cbind(ID,G)
CL2 <- rbind(ID,G)

CL1
CL2
```

11. Construct a matrix with predetermined dimensions:

```{r}
CL <- matrix(nrow=5, ncol=2)
CL[,1] <- ID
CL[,2] <- G
colnames(CL) <- c("ID", "Grades")
```

12. Find the dimension of a matrix:

```{r}
dim(CL)
```

13. Construct a vector with repeated elements:

```{r}
GG1 <- rep(G, 3)
GG2 <- rep(G, each=3)

GG1
GG2
```

14. Construct a vector with a sequence of elements:

```{r} 
b <- seq(from=1, to=10, by=2)

b
```

15. Remove a variable:

```{r}
rm(b)
exists("b")

```

16. Working with Normal distribution:

```{r}
dnorm(0, mean = 0, sd = 1)
pnorm(1.96, mean = 0, sd = 1)
qnorm(0.975, mean = 0, sd = 1)
SND <- rnorm(100, mean = 0, sd = 1)

SND
```

17. Shapiro Normality test:

```{r}
shapiro.test(SND)
```

18. Test of correlation:

```{r}
SND2 <- rnorm(100, mean = 0, sd = 1)
cor.test(SND,SND2)
```

19. Change the seed number:

```{r}
set.seed(164)
rnorm(1, 0, 1)
rnorm(1, 0, 1)
set.seed(164)
rnorm(1, 0, 1)
```

20. Generate Standard Normal random variate:

```{r}
u1 <- runif(10,0,1)
u2 <- runif(10,0,1)
z1 = sqrt(-2*log(u1))*cos(2*pi*u2)
z2 = sqrt(-2*log(u1))*sin(2*pi*u2)

z1
z2
```

21. Close R:

```{r, eval=FALSE}
q()
```