---
title: "Lab2"
author: "B. Moran"
date: "9 August 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Computer Lab. 2 {Week 3}


###1	Problems


1.	Question 1 - Assignment 1.

2.	Question 2 - Assignment 1.

3.	Bonus Question - Assignment 1.


###2	Random Variate Generation in R


1.	Congruence relationship $\implies a \equiv  b (\text{mod } n)$:
```{r congruence}
10 %% 3
```

2.	What is a random number?

**Answer** a random number generated using $Unif[0,1]$

3.	Change the seed number: 
```{r change seed}
set.seed(164) 
runif(1, 0, 1) 
runif(1, 0, 1) 
set.seed(164) 
runif(1, 0, 1)

```

4.	What is a random variate?

**Answer** 

5.	Let $X$ be a continuous random variable with the CDF $F_X(x)$. If we define the random variable $Y := F_X(X)$, show that $Y$ has the Uniform distribution over interval $[0, 1]$.

6.	Random variate generation from the Exponential distribution:
```{r exponential}

Sample <- -1/2*log(1-runif(1000,0,1)) 
M <- mean(Sample)

M
```

7.	Plot histogram: 
```{r plot}
hist(Sample, 50)
```

8.	Kolmogorov-Smirnov goodness-of- t test: 
```{r gof}
ks.test(Sample, "pexp", 1/M)
```

9.	How to generate random variates from distributions that the CDF can be calculated explicitly?

10.	Let $R_1$ and $R_2$ be two independent random variables having the common Uniform distribution over interval $[0, 1]$. Then, two random variables $Z_1$ and $Z_2$ defined as
$$\begin{aligned} Z_1 &:=	\sqrt{-2\log(R_1)}\cos(2\pi R_2) \\
Z2 &:=	\sqrt{-2\log(R_1)}\sin(2\pi R_2 \end{aligned},$$

are independent and identically distribution having the Standard Normal distribution.

