---
title: "lab3"
author: "Benjamin Moran"
date: "16 August 2016"
output: html_document
---

##R-Package: ASTSA
1. Download the supplementary R-package provided by the authors of the main reference book:
ASTSA: Applied Statistical Time Series Analysis.
This package can be accessed for free from http://www.stat.pitt.edu/stoffer/tsa4/.
2. Save the zipped package in a desired directory, say ‘C:/ASTSA’.
3. Set the current working directory where the package has been saved:
setwd("C:/ASTSA")
4. Install the package:
```{r install, eval=FALSE}
install.packages("astsa")
```
5. Load the package:
```{r load_astsa}
library(astsa)
```
6. The yearly average global temperature deviations (1880-2009) in degrees centigrade:
```{r gtem}
gtemp
```
7. Plot ‘gtemp’:
```{r gtemp}
plot(gtemp)

plot(gtemp, type="o")

?plot

plot(gtemp, type="o", ylab="Global Temperature Deviations")

plot(gtemp, type="o", ylab=expression(Global~Temperature~Deviations))

plot(gtemp, type="o", ylab=expression(bold(Global)~bold(Temperature)~bold(Deviations)))

plot(gtemp, type="o", main="Global Temperature Plot",
ylab=expression(bold(Global) ~ bold(Temperature) ~ bold(Deviations)))
```

8. Save the plot.

9. Set the Graphical Parameters:
```{r params}
?par
```

10. Plot the SOI and Recruitment time series together:
```{r plotsoi}
par(mfrow = c(2,1))
plot(soi, ylab="", xlab="", main="Southern Oscillation Index")
plot(rec, ylab="", xlab="", main="Recruitment")
par(mfrow = c(1,1))
plot(gtemp, type="o", main="Global Temperature Plot",
ylab=expression(bold(Global)~bold(Temperature)~bold(Deviations)))
```

11. Plotting ACF and CCF:
```{r plotacf}
par(mfrow = c(3,1))
acf(soi, 48, main="Southern Oscillation Index")
acf(rec, 48, main="Recruitment")
ccf(soi,rec, main="SOI vs. Recruitment", ylab="CCF")
```

12. Plotting autocovariance and cross-covariance functions:
```{r plotccf}
par(mfrow = c(3,1))
acf(soi, 48, type=c("covariance"), main="Southern Oscillation Index")
acf(rec, 48, type=c("covariance"), main="Recruitment")
ccf(soi,rec, type=c("covariance"), main="SOI vs. Recruitment", ylab="CCF")
```