---
title: "Assignment3R"
author: "Benjamin Moran"
date: "11 September 2016"
output: html_document
---
```{r Setup, include=FALSE}
require(broom)
require(knitr)
require(tidyr)
require(magrittr)
require(ggfortify)
require(ggplot2)
```


```{r Q7a, eval = FALSE}

setwd("/Users/benjamin/R/STAT3040/Data")

```

8.	Store the data from the  file to a variable in R:

```{r Q8a}
SF.US <- readRDS("GTSugarFatTS.rds")
```

9.	Look at the structure of the variable: 

```{r Q9a, echo=FALSE}

head(SF.US)

```

###Linear Regression in R


1.	Find the time range of dataset:

```{r Q1b}

SF.US[1,1]

dim(SF.US)

SF.US[152,1]

```
2.	Construct two time series for Sugar and Fat searches in the United States within the range of observation times:

```{r Q2b}

Sugar.TS <- ts(SF.US[,2], start = c(2004,1), end = c(2016,8), frequency = 12) 
Sugar.TS

Fat.TS <- ts(SF.US[,3], start = c(2004,1), end = c(2016,8), frequency = 12) 
Fat.TS

Fat.TS2 <- fortify(Fat.TS) %>% 
  separate(Index, c("Year", "Month", "Day")) %>%
  subset(select = -Day)

Sugar.TS2 <- fortify(Sugar.TS) %>% 
  separate(Index, c("Year", "Month", "Day")) %>%
  subset(select = -Day)
```


3.	Plot the time series of Sugar.TS and Fat.TS:

```{r Q3b}

plot.ts(Sugar.TS, type="l") 
plot.ts(Fat.TS, type="l") 
plot.ts(cbind(Sugar.TS, Fat.TS), type="l") 
ts.plot(cbind(Sugar.TS, Fat.TS), type="l")


```

```{r Q3b alt}
ggplot(Fat.TS2, aes(Month, Data, group = Year, colour = Year)) + 
  geom_line()

ggplot(Sugar.TS2, aes(Month, Data, group = Year, colour = Year)) + 
  geom_line()

```

4.	Construct a time variable for observation times:
```{r Q4b}
TimeTS <- time(Sugar.TS) 
TimeTS
```

5.	Plot the scatterplot matrix:

```{r Q5b}
data <- data.frame(Sugar = Sugar.TS, Time = TimeTS, Fat = Fat.TS)
pairs(cbind(Sugar = Sugar.TS, Time = TimeTS, Fat = Fat.TS))
```

6. Fit two regression models
$$
  \begin{cases}
    \begin{aligned}
      \text{Model 1: Sugar}_{t} &=	\beta_{1} +  \beta_{2}t +  \beta_{3}\text{Fat}_{t} + \mathcal{W}_{t} \\
       \text{Model 2: Sugar}_{t} &=	\beta_{1} +  \beta_{2}t +  \beta_{3}t^{2} +  \beta_{4}\text{Fat}_{t} + \mathcal{W}_{t}
    \end{aligned}
  \end{cases}
$$
```{r Q6b}
fit1 <- lm(Sugar.TS ~ TimeTS + Fat.TS) 
TimeTS2 <- TimeTS^2
fit2 <- lm(Sugar.TS ~ TimeTS + TimeTS2 + Fat.TS)

```

7.	Summary of the regression parameters estimations: 

```{r Q7b}
glance(fit1)
tidy(fit1)
glance(fit2)
tidy(fit2)
```


8.	Compare the values of AIC for two models:

```{r Q8b}
AIC(fit1)/152 - log(2*pi)
AIC(fit2)/152 - log(2*pi)
```

