---
title: "STAT3040 : Time Series Analysis 2016 Assignment 5"
output: html_notebook
---

```{r setup, include = FALSE}
require(broom)
require(astsa)
require(knitr)
require(tidyr)
require(magrittr)
require(ggfortify)
require(ggplot2)
require(ggthemes)
knitr::opts_chunk$set(echo = FALSE, collapse = TRUE, fig.align="center", fig.width=15, fig.height = 7.5)
options(digits = 16)
```

### Submission:
You should email your R-code to `Ali.Eshragh@newcastle.edu.au' by Sunday, October 23, 11:55pm.  Write your full name and student ID No.  on the top of your code and your  file name must be your complete surname.

#### 1.  Find  the  theoretical  autocorrelation  function  of  the   first  AR  model  given  on  the  top  of  Slide 40, Lecture Slides Chapter 3, explicitly, and compare its value in lag 5 with the corresponding output of ARMAacf.

The model is:

$$
X_t= 0.7X_{t-1} - 0.12 X_{t-2} + \mathcal{W}_{t}
$$
First, we find the roots of the model using polyroot.
```{r Ass51}
z <- c(1,-0.7,0.12)
Roots <- polyroot(z)
Roots <- Re(Roots)
Roots
```

```{r Ass52}
A <- matrix(nrow = 2, ncol = 2)
A[,1] <- c(1,Roots[1])
A[,2] <- c(1,Roots[2])
B <- c(1, 1/1.12)
Sol <- solve(A,B)
Sol
```
```{r Ass53}
Sol[1]*Roots[1]^(-5) + Sol[2]*Roots[2]^(-5)
ACF4AR2 <- ARMAacf(ar=c(0.7,-0.12), ma=0, 100)
ACF4AR2[6]
```

2.  Plot the theoretical and sample PACF of all three AR models given in Slide 40, Lecture Slides Chapter 3.
$$
X_t= 0.7X_{t-1} - 0.12 X_{t-2} + \mathcal{W}_{t} \\
X_t= X_{t-1} - 0.25 X_{t-2} + \mathcal{W}_{t} \\
X_t= 1.5X_{t-1} - 0.75 X_{t-2} + \mathcal{W}_{t}
$$

```{r Ass5Q2}
PACF4AR2A <- ARMAacf(ar=c(0.7,-0.12), ma=0, 50, pacf=TRUE)
plot(PACF4AR2A, type="h")
abline(h=0)
AR2A <- arima.sim(list(order=c(2,0,0), ar=c(0.7,-0.12)), n=1000)
pacf(AR2A, 50)


PACF4AR2B <- ARMAacf(ar=c(1,-0.25), ma=0, 50, pacf=TRUE)
plot(PACF4AR2B, type="h")
abline(h=0)
AR2B <- arima.sim(list(order=c(2,0,0), ar=c(1,-0.25)), n=1000)
pacf(AR2B, 50)


PACF4AR2C <- ARMAacf(ar=c(1.5,-0.75), ma=0, 50, pacf=TRUE)
plot(PACF4AR2C, type="h")
abline(h=0)
AR2C <- arima.sim(list(order=c(2,0,0), ar=c(1.5,-0.75)), n=1000)
pacf(AR2C, 50)
```