---
title: "Assignment3R"
author: "Benjamin Moran"
date: "1 September 2016"
output: html_document
---

```{r Setup, include=FALSE}
require(broom)
require(knitr)
require(tidyr)
require(magrittr)
require(ggfortify)
require(ggplot2)
```

###Getting Started the Project


1.	Open the page https://www.google.com.au/trends/.

2.	Type a search keyword, say `wine'.

3.	Choose a country, say `Australia'.

4.	Choose a compare search keyword, say `beer'.

5.	Save the data  file with extension .csv' on your computer, say test.csv'.

6.	Rearrange your data  file such that each column contains the heading and data only.

7.	Set your current working directory in R where the le has been saved.

```{r Q7a}

setwd("/Users/benjamin/R/STAT3040/Assignments")

```

8.	Store the data from the  le to a variable in R:

```{r Q8a}
WBAust <- read.csv("GTWineBeerTs.csv", header=TRUE, skip = 1)
```

9.	Look at the structure of the variable: 

```{r Q9a, echo=FALSE}

head(WBAust)

```

10.	All students should form groups of size 1-3 to carry out the project. The list of each group mem-bers should be sent to `Ali.Eshragh@newcastle.edu.au' by Wednesday, September 7, 11:55pm.

11.	Those students that do not join a group, should attend the consultation session in V239 on Thursday, September 8, 11:30am-12:30pm to discuss the possibility of forming new groups to-gether.

12.	Each group should send their suggested data le with a short description to `Ali.Eshragh@newcastle.edu.au' by Wednesday, October 5, 11:55pm. You are free to nd your dataset from any permitted re-

source. `Google-Trend' is only one suggestion.

###Linear Regression in R


1.	Find the time range of dataset:

```{r Q1b}

WBAust[1,1]

dim(WBAust)

WBAust[152,1]

```
2.	Construct two time series for wine and beer consumption in Australia within the range of ob-servation times:

```{r Q2b}

WineTS <- ts(WBAust[,2], start = c(2004,1), end = c(2016,8), frequency = 12) 
WineTS

BeerTS <- ts(WBAust[,3], start = c(2004,1), end = c(2016,8), frequency = 12) 
BeerTS
BeerTS2 <- fortify(BeerTS) %>% 
  separate(Index, c("Year", "Month", "Day")) %>%
  subset(select = -Day)
```


3.	Plot the time series of WineTS and BeerTS:

```{r Q3b}

plot.ts(WineTS, type="l") 
plot.ts(BeerTS, type="l") 
plot.ts(cbind(WineTS, BeerTS), type="l") 
ts.plot(cbind(WineTS, BeerTS), type="l")

```

```{r Q3b alt}
ggplot(BeerTS2, aes(Month, Data, group = Year, colour = Year)) + 
  geom_line()

```

4.	Construct a time variable for observation times:
```{r Q4b}
TimeTS <- time(WineTS) 
TimeTS
```

5.	Plot the scatterplot matrix:

```{r Q5b}
data <- data.frame(Wine = WineTS, Time = TimeTS, Beer = BeerTS)
pairs(cbind(Wine = WineTS, Time = TimeTS, Beer = BeerTS))
```

6. Fit two regression models
$$
  \begin{cases}
    \begin{aligned}
      \text{Model 1: Wine}_{t} &=	\beta_{1} +  \beta_{2}t +  \beta_{3}\text{beer}_{t} + \mathcal{W}_{t} \\
       \text{Model 2: Wine}_{t} &=	\beta_{1} +  \beta_{2}t +  \beta_{3}t^{2} +  \beta_{4}\text{beer}_{t} + \mathcal{W}_{t}
    \end{aligned}
  \end{cases}
$$
```{r Q6b}
fit1 <- lm(WineTS ~ TimeTS + BeerTS) 
TimeTS2 <- TimeTS^2
fit2 <- lm(WineTS ~ TimeTS + TimeTS2 + BeerTS)

```

7.	Summary of the regression parameters estimations: 

```{r Q7b}
glance(fit1)
tidy(fit1)
glance(fit2)
tidy(fit2)
```


8.	Compare the values of AIC for two models:

```{r Q8b}
AIC(fit1)/152 - log(2*pi)
AIC(fit2)/152 - log(2*pi)
```





####References

[1]	R.H. Shumway and D.S. Sto er, Time Series Analysis and Its Applications With R Examples, Springer, New York, 2010.

