---
title: 'STAT3040 : Time Series Analysis 2016 Lab 11'
output:
  html_document: default
  html_notebook: default
---

```{r setup, include = FALSE}
require(broom)
require(astsa)
require(knitr)
require(tidyr)
require(magrittr)
require(ggfortify)
require(ggplot2)
require(ggthemes)
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, fig.align="center", fig.width=15, fig.height = 7.5)
```

### Pre-Whitening


1.	The basic problem that were considering is the construction of a lagged regression in which we predict a $Y_t$ variable at time $t$ using lags of an $X_t$ variable (including lag $0$) and lags of the $Y_t$ variable.

2.	We have already introduced the cross-correlation function as an aid to the identification of the model.

3.	One dificulty is that the CCF is affected by the time series structure of the $X_t$ variable and any `in common' trends the time series $X_t$ and $Y_t$ may have over time.

4.	One strategy for dealing with this difficulty is called pre-whitening. The steps are:

    (a)	Determine a time series model for the $X_t$  variable and store the residuals from this model.
    (b)	Filter the $Y_t$ variable series using the $X_t$ variable model (using the estimated coe cients from step (a)). In this step we find differences between observed $Y_t$ values and 'estimated' $Y_t$ values based on the $X_t$ variable model.
    (c)	Examine the CCF between the residuals from Step (a) and the  filtered $Y_t$  values from Step (b). This CCF can be used to identify the possible terms for a lagged regression.

5.	Generate a time series from the AR(1) model $X_t = 0.8X_{t-1} + \mathcal{W}_t$: 
```{r Q5}
x <- arima.sim(list(c(1,0,0), ar=0.8), n=200)
```

6.	Generate the time series $Y_t = 10 + 0.6X_{t-2} + 1.2X_{t-4} + \mathcal{V}_t$ where $\mathcal{V}_t$ is a Guassian White Noise series with mean zero and variance one:
```{r Q6}
z <- ts.intersect(x, lag(x,-2), lag(x,-4)) 
dim(z)
y <- 10 + 0.6*z[,2] + 1.2*z[,3] + rnorm(196,0,1)
```

7.	Fit an ARIMA model to $X_t$: 
```{r Q7}
acf(x)
pacf(x)
fit <- arima(x, c(1,0,0)) 
fit$coef
```

8.	Filter the time series $Y_t$  using the model for $X_t$:
```{r Q8}
pwy <- filter(y, filter = c(1,-fit$coef[1]), sides = 1)
head(pwy)
```

9.	Plot the cross-correlation function between the residuals of the tted model for the time series $X_t$ and the filtered model for $Y_t$:
```{r Q9}
pwx <- fit$residuals 
ccf(pwx,pwy, na.action=na.omit)
```
10.	Fit a lagged-regression model for the time series $Y_t$:
```{r Q10}
fit2 <- lm(y~z[,2]+z[,3]) 
summary(fit2)

fit3 <- arima(y, c(0,0,0), xreg = cbind(z[,2], z[,3])) 
fit3$coef
```