---
title: "lab4"
author: "B. Moran"
date: "24 August 2016"
output: 
  html_document:
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, highlight = TRUE)
require(astsa)
require(knitr)
require(broom)
require(ggplot2)
require(ggfortify)
```
###Generating Synthetic Time Series in R

1.	Create the vector $u = [1, 2, 3, 4, 5, 6,	7, 8, 9 ]'$:												
```{r Q1}
u <- seq(1,9, by =1)												
```

2.	Generate the moving average series $v_{t}$ according to $v_{t}	= \frac{1}{2}u_{t}+ \frac{1}{3}u_{t-1} + \frac{1}{6}u_{t-2}$ for $t = 3, 4, \cdots ,9$:	
```{r Q2}
v <- filter(u, sides = 1, c(1/2, 1/3, 1/6), method = ("convolution"))	
v													
v2 <- filter(u, sides = 1, c(1/2,1/3,1/6))												
v2
```

3.	Generate the moving average series $w_{t}$	according to $w_{t}	= \frac{1}{2}u_{t+1} + \frac{1}{3}u_{t} + \frac{1}{6}u_{t-1}$ for $t = 2, 3, \cdots ,8$:	
```{r Q3}
w <- filter(u, sides = 2, c(1/2,1/3,1/6), method = ("convolution")) 
w
w2 <- filter(u, c(1/2,1/3,1/6))
w2
```

4.	What will happen if the length of filter is an even number? Run the following code and see the output:

```{r Q4}
x <- filter(u, rep(1,4))
x
```

5.	Generate the autoregressive series yt according to $y_{t}	= -2y_{t-1} + 3y_{t-2} + u_{t}$ for $t = 1,2,\cdots ,9$	 with the initial values $y_{0} = -1$ and $y_{1} = 4$:

```{r Q5}
y <- filter(u, c(-2, 3), method = ("recursive"), init = c(-1,4))
y
```

6.	Generate the autoregressive series zt according to $z_{t}	= -2z_{t-1} + 3z_{t-2} + u_{t}$ for $t = 1,2,\cdots ,9$ with the initial values $z_{0} = z_{1} = 0$:

```{r Q6}
z <- filter(u, c(-2, 3), method = ("recursive"), init = c(0,0)) 
z
z <- filter(u, c(-2, 3), method = ("recursive")) 
z
```

7.	(a) Generate $n = 100$ observations from the autoregressive model

$$X_{t} =	-0.9X_{t-2} + \mathcal{W}_{t}$$

where $\mathcal{W}_{t}$ is Gaussian White Noise with $\sigma_\mathcal{W}^{2} = 1$.

```{r Q7a}
w = rnorm(100,0,1)
x <- filter(w, c(0,-0.9), method = "recursive")
x.f <- fortify(x)
```

(b) Next, generate the moving average series

$$Y_{t} = \frac{X_{t} + X_{t-1} + X_{t-2} + X_{t-3}}{4}$$

```{r Q7b}
y <- filter(x, sides = 1, rep(1/4,4))
y.f <- fortify(y)
```

(c)	Plot $X_{t}$ as a line and superimpose $Y_{t}$ as a dashed line. 

```{r Q7c}
plot(x, type="l")
lines(y, lty = "dashed")
```
```{r Q7cggplot}
ggplot() +
  geom_line(data = x.f, aes(Index, Data), colour = "Blue") +
  geom_line(data = y.f, aes(Index, Data), colour = "Green", linetype = 2) + 
  theme_light()
```

(d)	Plot the ACF of $X_{t}$ and $Y_{t}$ next to each other. 

```{r Q7d}
par(mfrow=c(2,1))

autoplot(acf(x))

autoplot(acf(y, na.action = na.pass))
```

####References

R.H. Shumway and D.S. Sto er, Time Series Analysis and Its Applications With R Examples, Springer, New York, 2010.
