---
title: 'STAT3040 : Time Series Analysis 2016 Computer Lab. 9 - Week 10'
output:
  html_notebook: default
  html_document: default
---

```{r setup, include = FALSE}
require(broom)
require(astsa)
require(knitr)
require(tidyr)
require(magrittr)
require(ggfortify)
require(ggplot2)
require(ggthemes)
knitr::opts_chunk$set(echo = FALSE, collapse = TRUE, fig.align="center")
```

### 1.  ARMA Models in R

1.  Generate a time series from the MA(3) model
$$
X_t=\mathcal{W}_t + 0.8\mathcal{W}_{t-1} + 0.1\mathcal{W}_{t-2} - 0.7\mathcal{W}_{t-3} 
$$
```{r ARMA1}
MA3 <- arima.sim(list(order=c(0,0,3), ma=c(0.8,0.1,-0.7)), n=1000)
acf(MA3)
pacf(MA3)
```
2.  Generate a time series from the AR(2) model
$$
X_t=X_{t-1} - 0.16 X_{t-2} + \mathcal{W}_{t} 
$$
```{r ARMA2}
AR2 <- arima.sim(list(order=c(2,0,0), ar=c(1,-0.16)), n=1000)
acf(AR2, 50)
pacf(AR2, 50)
```

3.  Calculate the theoretical ACF and PACF of the AR(2) model
$$
X_t=X_{t-1} - 0.16 X_{t-2} + \mathcal{W}_{t} 
$$
```{r ARMA3}
ACF4AR2 <- ARMAacf(ar=c(1,-0.16), ma=0, 50)
plot(ACF4AR2, type="h")
abline(h=0)
PACF4AR2 <- ARMAacf(ar=c(1,-0.16), ma=0, 50, pacf=TRUE)
plot(PACF4AR2, type="h")
abline(h=0)
```

4.  Find the roots of the AR polynomial for the AR(2) model
$$
X_t=X_{t-1} - 0.16 X_{t-2} + \mathcal{W}_{t} 
$$
```{r ARMA4}
z <- c(1,-1,0.16)
Roots <- Re(polyroot(z))
Roots
```

5.  Find the theoretical autocorrelation function of the AR(2) model
$$
X_t=X_{t-1} - 0.16 X_{t-2} + \mathcal{W}_{t} 
$$
explicitly.
```{r ARMA5}
A <- matrix(nrow = 2, ncol = 2)
A[,1] <- c(1,Roots[1])
A[,2] <- c(1,Roots[2])
B <- c(1, 1/1.16)
Sol <- solve(A,B)
Sol
```

6. Based on the results of Item 5, calculate  the ACF in lags 1 and 12 and  compare them  with ACF4AR2:

```{r ARMA6}
# lag 1
Sol[1]*Roots[1]^(-1)+Sol[2]*Roots[2]^(-1)
# lag -1
Sol[1]*1.25^(1)+Sol[2]*5^(1)
# lag 1
ACF4AR2[2]
# lag 12
Sol[1]*1.25^(-12)+Sol[2]*5^(-12)
ACF4AR2[13]
```

7.  Load the 'ASTSA' package.
```{r ARMA7, eval=FALSE}
require(astsa)  
```

8.  Plot the Recruitment time series:
```{r ARMA8}
plot.ts(rec)
```

9.  Plot the sample ACF and PACF of the Recruitment time series in 25 lags:
```{r ARMA9}
acf(rec,25)
pacf(rec,25)
```

10.  Fit an AR(2) to the Recruitment time series:
```{r ARMA10}
fit <- arima(rec, c(2,0,0))
fit$coef
fit2 <- arima(rec, c(2,0,0), include.mean = FALSE)
fit2$coef
fit$aic
fit2$aic
```

11.  Calculate the p-value of the estimated coefficients:
```{r ARMA11}
sarima(rec,2,0,0)
```

12.  Analyse the residuals:
```{r ARMA12}
tsdiag(fit)
```

13.  Forecast the next three values of the time series:
```{r ARMA13, fig.height=15}
predict(fit, 3)
```

14.  Over fitting:

```{r ARMA14a, fig.height=15}
sarima(rec, 3,0,0)
```
```{r ARMA14b, fig.height=15}
sarima(rec, 2,0,1)
```

15.  Plot the time series vs. the fitted model:
```{r ARMA15}
RecData <- ts.intersect(rec, rec1=lag(rec,-1), rec2=lag(rec,-2))
fit <- lm(rec~rec1+rec2, data=RecData, na.action=NULL)
summary(fit)
fit2 <- lm(rec[3:453]~rec[2:452]+rec[1:451])
summary(fit2)
ts.plot(cbind(rec[3:453], fitted(fit)), col=c(1,4))
```



### References
R.H. Shumway and D.S. Stoffer,
Time Series Analysis and Its Applications With R Examples, Springer, New York, 2010.
