---
title: "STAT3040 : Time Series Analysis 2016 Computer Lab. 9 - Week 10"
output: html_notebook
---

```{r setup, include = FALSE}
require(broom)
require(astsa)
require(knitr)
require(tidyr)
require(magrittr)
require(ggfortify)
require(ggplot2)
require(ggthemes)
knitr::opts_chunk$set(echo = FALSE, collapse = TRUE, fig.align="center", fig.width=15, fig.height = 7.5)
```

### 1.  ARMA Models in R

1.  Generate a time series from the MA(3) model
$$
X_t=\mathcal{W}_t + 0.8\mathcal{W}_{t-1} + 0.1\mathcal{W}_{t-2} - 0.7\mathcal{W}_{t-3} 
$$
```{r ARMA1}
MA3 <- arima.sim(list(order=c(0,0,3), ma=c(0.8,0.1,-0.7)), n=1000)
acf(MA3)
pacf(MA3)
```
2.  Generate a time series from the AR(2) model
$$
X_t=X_{t-1} - 0.16 X_{t-2} + \mathcal{W}_{t} 
$$
```{r ARMA2}
AR2 <- arima.sim(list(order=c(2,0,0), ar=c(1,-0.16)), n=1000)
acf(AR2, 50)
pacf(AR2, 50)
```

3.  Calculate the theoretical ACF and PACF of the AR(2) model
$$
X_t=X_{t-1} - 0.16 X_{t-2} + \mathcal{W}_{t} 
$$
```{r ARMA3}
ACF4AR2 <- ARMAacf(ar=c(1,-0.16), ma=0, 50)
plot(ACF4AR2, type="h")
abline(h=0)
PACF4AR2 <- ARMAacf(ar=c(1,-0.16), ma=0, 50, pacf=TRUE)
plot(PACF4AR2, type="h")
abline(h=0)
```

4.  Find the roots of the AR polynomial for the AR(2) model
$$
X_t=X_{t-1} - 0.16 X_{t-2} + \mathcal{W}_{t} 
$$
```{r ARMA4}
z <- c(1,-1,0.16)
Roots <- polyroot(z)
Roots
```

5.  Find the theoretical autocorrelation function of the AR(2) model
$$
X_t=X_{t-1} - 0.16 X_{t-2} + \mathcal{W}_{t} 
$$
explicitly.
```{r ARMA5}
A <- matrix(nrow = 2, ncol = 2)
A[,1] <- c(1,1.25)
A[,2] <- c(1,5)
B <- c(1, 1/1.16)
Sol <- solve(A,B)
Sol
```

6. Based on the results of Item 5, calculate  the ACF in lags 1 and 12 and  compare them  with ACF4AR2:

```{r ARMA6}
Sol[1]*1.25^(-1)+Sol[2]*5^(-1)
ACF4AR2[2]
Sol[1]*1.25^(-12)+Sol[2]*5^(-12)
ACF4AR2[13]
```

7.  Load the 'ASTSA' package.
```{r ARMA7, eval=FALSE}
require(astsa)  
```

8.  Plot the Recruitment time series:
```{r ARMA8}
plot.ts(rec)
```

9.  Plot the sample ACF and PACF of the Recruitment time series in 25 lags:
```{r ARMA9}
acf(rec,25)
pacf(rec,25)
```

10.  Fit an AR(2) to the Recruitment time series:
```{r ARMA10}
fit <- arima(rec, c(2,0,0))
fit$coef
fit2 <- arima(rec, c(2,0,0), include.mean = FALSE)
fit2$coef
fit$aic
fit2$aic
```

11.  Calculate the p-value of the estimated coefficients:
```{r ARMA11}
sarima(rec,2,0,0)
```

12.  Analyse the residuals:
```{r ARMA12}
tsdiag(fit)
```

13.  Forecast the next three values of the time series:
```{r ARMA13, fig.height=15}
predict(fit, 3)
```

14.  Over fitting:

```{r ARMA14a, fig.height=15}
sarima(rec, 3,0,0)
```
```{r ARMA14b, fig.height=15}
sarima(rec, 2,0,1)
```

15.  Plot the time series vs. the fitted model:
```{r ARMA15}
RecData <- ts.intersect(rec, rec1=lag(rec,-1), rec2=lag(rec,-2))
fit <- lm(rec~rec1+rec2, data=RecData, na.action=NULL)
summary(fit)
fit2 <- lm(rec[3:453]~rec[2:452]+rec[1:451])
summary(fit2)
ts.plot(cbind(rec[3:453], fitted(fit)), col=c(1,4))
```

### 2  Assignment 5: R Coding
#### Submission:
You should email your R-code to `Ali.Eshragh@newcastle.edu.au' by Sunday, October 23, 11:55pm.  Write your full name and student ID No.  on the top of your code and your  file name must be your complete surname.

1.  Find  the  theoretical  autocorrelation  function  of  the   first  AR  model  given  on  the  top  of  Slide 40, Lecture Slides Chapter 3, explicitly, and compare its value in lag 5 with the corresponding output of ARMAacf.

The three models are:

$$
X_t= 0.7X_{t-1} - 0.12 X_{t-2} + \mathcal{W}_{t}
$$
```{r Ass51}
z <- c(1,-0.7,0.12)
Roots <- polyroot(z)
Roots
```

```{r Ass52}
A <- matrix(nrow = 2, ncol = 2)
A[,1] <- c(1,2.5)
A[,2] <- c(1,10/3)
B <- c(1, 1/1.12)
Sol <- solve(A,B)
Sol
```
```{r Ass53}
Sol[1]*2.5^(-5)+Sol[2]*(10/3)^(-5)

ACF4AR2[6]
```

2.  Plot the theoretical and sample PACF of all three AR models given in Slide 40, Lecture Slides Chapter 3.
$$
X_t= 0.7X_{t-1} - 0.12 X_{t-2} + \mathcal{W}_{t} \\
X_t= X_{t-1} - 0.25 X_{t-2} + \mathcal{W}_{t} \\
X_t= 1.5X_{t-1} - 0.75 X_{t-2} + \mathcal{W}_{t}
$$

```{r Ass5Q2}
PACF4AR2 <- ARMAacf(ar=c(0.7,-0.12), ma=0, 50, pacf=TRUE)
plot(PACF4AR2, type="h")
abline(h=0)
AR2 <- arima.sim(list(order=c(2,0,0), ar=c(0.7,-0.12)), n=1000)
pacf(AR2, 50)


PACF4AR2 <- ARMAacf(ar=c(1,-0.25), ma=0, 50, pacf=TRUE)
plot(PACF4AR2, type="h")
abline(h=0)
AR2 <- arima.sim(list(order=c(2,0,0), ar=c(1,-0.25)), n=1000)
pacf(AR2, 50)


PACF4AR2 <- ARMAacf(ar=c(1.5,-0.75), ma=0, 50, pacf=TRUE)
plot(PACF4AR2, type="h")
abline(h=0)
AR2 <- arima.sim(list(order=c(2,0,0), ar=c(1.5,-0.75)), n=1000)
pacf(AR2, 50)
```

### References
R.H. Shumway and D.S. Stoffer,
Time Series Analysis and Its Applications With R Examples, Springer, New York, 2010.
