---
title: Thorough research of this special topic regarding the influence of various factors
author: [N. Della, T. McCallum, B. Moran]
institute: School of Science & I.T., UoN
longinstitute: School of Science and Information Technology, University of Newcastle
web: 'newcastle.edu.au'
posteroptions: width=90,height=110,scale=1.2 #,grid
headerheight: 13cm
# large, Large, LARGE, huge, Huge, veryHuge, VeryHuge, VERYHuge
titlefont: size=\veryHuge,series=\bfseries
authorfont: size=\huge
institutefont: size=\Large
knit: (function(input, encoding, make = TRUE) { source('templates/makefile-renderer.R', local =TRUE) })
---
%% smart
%% to=latex
%% template=templates/poster.tex
%% filter=templates/poster-filters.py

```{r packages-and-options, cache=FALSE}
library('ggplot2')
library('knitr')
theme_set(theme_grey(base_size=10))
knitr::opts_chunk$set(fig.width = 16, fig.height = 4)
library(astsa)
file.loc <- "fludata3.csv"
fludata <- read.csv(file.loc, header=TRUE)
CItestfunc <- function( # Calcs if 95% of acf/pacf within blue CI bars.
  x,
  cnf.int = qnorm({1 + 0.95}/2)/sqrt(as.integer(length(x)))) { 
ans <- mean(abs(x)<cnf.int)
ans2 <- ifelse(ans<0.95,"No","Yes")
print(ans)
print(ans2)
}
```
\vskip 3cm
[columns = 2]
[column]
Two lines adfaksjd;gadg;skljf;alsdjf;lasdjflkasdl;f afkaskjdf asjdfk alksdjfal; ksdjf ;ajsdf a;lksdjf ;alksjdf ;lakjsdf ;aklsjdf ;lakjsdf akjsd lfkjasdjfajkshgaj; lsdkgjoaidgsha osdljfn aow;ehjfnapoweu sdvn;sodlk v;alkcxnva ;sldnaosd;lksand;lkf jva;mdskvj aosdijg  ds alskjdfl asdj ll  kalsdjf a;sldf as;dlf asdklfasdlkf asdfklasdfjdkf a;lsdkfj alsdjf l;asdlkfjasld;f asjdf dsal;fkdsjf asdjfla ;sdjf asdjfl .

```{r test2}
x <- fludata$ILI_Unvaccinated
pr.ix <- which.min(x)
x[pr.ix] <- {x[pr.ix-3] + 2*x[pr.ix-2] + 3*x[pr.ix-1] + 3*x[pr.ix+1] + 2*x[pr.ix+2] + x[pr.ix+3]}/12
plot.ts(x)
```



Two lines adfaksjd;gadg;skljf;alsdjf;lasdjflkasdl;f afkaskjdf asjdfk alksdjfal; ksdjf ;ajsdf a;lksdjf ;alksjdf ;lakjsdf ;aklsjdf ;lakjsdf akjsd lfkjasdjfajkshgaj; lsdkgjoaidgsha osdljfn aow;ehjfnapoweu sdvn;sodlk v;alkcxnva ;sldnaosd;lksand;lkf jva;mdskvj aosdijg  ds alskjdfl asdj ll  kalsdjf a;sldf as;dlf asdklfasdlkf asdfklasdfjdkf a;lsdkfj alsdjf l;asdlkfjasld;f asjdf dsal;fkdsjf asdjfla ;sdjf asdjfl .
[/column]
[/columns]